---
title: "Immune Gene Expression in Barn Owls"
author: "Prathmesh Bhagwat"
date: "2025-04-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Understanding gene expression patterns in birds is essential for investigating host immune responses to infections. My master's thesis focuses on avian immune gene expression in barn owls (Tyto alba) co-infected with Leucocytozoon parasites and Matryoshka viruses. For this Biometry project, I am analyzing real RNA-seq gene count data generated by Caroline Faircloth’s project on barn owls, using simulated infection metadata to explore differential immune gene expression.

This analysis allows me to practice transcriptomic workflows and statistical techniques relevant to my thesis while evaluating gene expression variation across simulated infection status.

## Research Question & Aims

### Research Question:
How does immune gene expression vary between barn owls simulated as virus-infected and uninfected?

### Aim 1:
Load and explore RNA-seq gene expression data from 6 barn owl individuals.

### Aim 2:
Simulate virus infection status and identify differentially expressed genes between infected and uninfected groups.

## Hypothesis

**Null Hypothesis (H0):** There is no significant difference in immune gene expression between virus-infected and uninfected barn owls.

**Alternative Hypothesis (H1):** Immune gene expression significantly differs between virus-infected and uninfected barn owls.

```{r load-data}
# Load gene expression data from Caroline's dataset
gene_data <- read.csv("reads_per_gene/gene_expression_data.csv")
head(gene_data)

```
