---
title: "Immune Gene Expression in Barn Owls Co-infected with Leucocytozoon and Matryoshka Virus"
author: "Prathmesh Bhagwat"
date: "2025-04-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Understanding gene expression patterns in birds is essential for investigating host immune responses to infections. My master's thesis focuses on avian immune gene expression in barn owls (Tyto alba) co-infected with Leucocytozoon parasites and Matryoshka viruses. This co-infection model allows us to study synergistic or antagonistic effects of two different pathogens on host immune gene expression. For this Biometry project, I am analyzing real RNA-seq gene count data generated by Caroline Faircloth's project on barn owls, using simulated infection metadata to explore differential immune gene expression.

## Research Question & Aims

### Research Question:
How does immune gene expression vary between barn owls simulated as co-infected with Leucocytozoon and Matryoshka virus, versus uninfected?

### Aim 1:
Load and explore RNA-seq gene expression data from 6 barn owl individuals.

### Aim 2:
Simulate co-infection status and identify differentially expressed genes between co-infected and uninfected groups.

## Hypothesis
**Null Hypothesis (H0):** There is no significant difference in immune gene expression between barn owls co-infected with Leucocytozoon and Matryoshka virus and those uninfected.

**Alternative Hypothesis (H1):** Immune gene expression significantly differs between co-infected and uninfected barn owls.

```{r load-data}
# Load gene expression data from Caroline's dataset
gene_data <- read.csv("reads_per_gene/gene_expression_data.csv")
head(gene_data)

```
