---
title: "Biometry Intro"
author: "Prathmesh Bhagwat"
date: "2025-04-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Understanding gene expression in parasites is crucial for studying host-pathogen interactions. My thesis project investigates immune gene expression in birds co-infected with Leucocytozoon parasites and Matryoshka viruses. However, for this Biometry assignment, I am using a public RNA-seq dataset on Babesia bovis as a placeholder to simulate the analysis workflow.

This substitution allows me to practice working with metadata, loading data in R, and preparing for future transcriptomic analysis using avian samples.


## Research Question & Aims

### Research Question:
How does gene expression vary across two life stages of Babesia bovis (merozoite vs. kinete)?

### Aim 1:
Load and explore RNA-seq sample metadata for 6 samples of Babesia bovis.

### Aim 2:
Simulate differential gene expression comparison between merozoite and kinete stages using metadata.

```{r load-data}
# Load gene expression data from Caroline's dataset
gene_data <- read.csv("reads_per_gene/gene_expression_data.csv")
head(gene_data)

```